Binary Image Transition State Search Method

### 2022-03-30 manuscript.tex
Updated captions of figures

\documentclass[aps,prl,twocolumn,10pt,groupedaddress]{revtex4-2}

@@

%\section{Introduction \& Method}
Transition states are central to the description of reconfiguration mechanisms for systems in chemistry, condensed matter physics, and engineering.
Historically, many computational methods for locating transition states have grown from an atomistic or particulate perspective.
These have proven to be important tools for understanding, for example, protein folding \cite{Bryngelson1995,Onuchic1997}, biological and industrial catalysis \cite{Boehr2006,Kerns2015,Guo2018a}, quantum tunneling \cite{Richardson2016,Vaillant2019}, crystallisation \cite{Richard2018}, and cluster formation \cite{Wales1998,Wales2012}.

@@

%\section{Method}
The method begins by first initialising the states, $\bm{x}_1$ and $\bm{x}_2$, in the basins of attraction of different local minima.
These can be set to the minima, but this is not a necessary requirement.
The BITSS method then leads these states to meet at the lowest point on the ridge between the basins of attraction - the transition state, as shown in \cref{fig:toy2d}a.
This is done by minimising the energy of the pair while constraining their separation and energy difference.
The separation constraint $d_i$ is incrementally reduced, such that at iteration $i$,
\begin{equation}\label{eq:diteration}
  d_i = (1 - f) d_{i-1},
\end{equation}
with $d_0$ the separation between the initial states.
A factor of $f = 0.5$ is succesful for most applications, but this can be made smaller to ensure that the states do not slide off the ridge.
The secondary constraint sets the energies of the two states equal to ensure that neither state is pulled over the ridge.

@@

%\section{Speed Comparison}
\begin{figure*}[tb]
  \includegraphics{speedtest.pdf}
  \caption{\label{fig:speedtest}
    The convergence to the transition state as a function of the number of gradient calculations using the BITSS (black line) and string methods for (a) a Lennard-Jones seven-particle cluster, (b) cylindrical shell buckling, and (c) wetting of a chemically-striped surface.
    The string method is repeated with a differing number of images along the string, the number of which is listed in the legend in (c).
    The shown configurations correspond to the two minimum energy states and the transition state, marked by an asterisk.
  }
\end{figure*}
As a result of the BITSS method using only two states it has the potential to be significantly faster than chain-of-states approaches which typically use tens of states.
This is tested here by measuring the speed of convergence to the transition state for BITSS in comparison with the string method and DNEB for several representative examples.

@@

%\section{Adaptive mesh}
\begin{figure}[tb]
  \includegraphics{adaptivemesh.pdf}
  \caption{\label{fig:adaptivemesh}
    (a) Snapshots of the BITSS method applied to the buckling of a cylinder with a changing mesh.
        The radial displacement relative the unbuckled cylinder is shown, as well as the underlying triangular mesh.
    (b) BITSS applied to the striped wetting example with different resolutions for the two states.
        Each grid cell denotes 50x50 lattice nodes.
        The zoomed axis shows the difference in the fluid interface between the two final states, as well as the approximated transition state (solid black line).
        This is compared to the transition state found using a high resolution (dashed line).
  }
\end{figure}

@@

%\section{Zero-curvature modes}
\begin{figure}[tb]
  \includegraphics{flatdiscontinuous.pdf}
  \caption{\label{fig:flatdiscontinuous}
    (a) Energy profile of the BITSS pathway on a 2D potential with flat regions.
        Blue and red dots denote the minima and transition state, respectively.
        Points of interest are labelled by i--iii (see text).
        Top inset: A zoomed in view around the transition state.
        Bottom inset: The pathway taken, with the edges of the flat regions marked by dashed lines.
    (b) Disconnectivity graph of the energy landscape for a seven-particle cluster with a discontinuous hard-core pair potential.
    (c) Plot of the discontinuous potential.
        The repulsive section of the standard Lennard-Jones potential is also shown in grey.
  }
\end{figure}

@@

%\section{Gradient-free problems}
Finally, we investigate the application of BITSS to systems with undefined gradients, such as when the landscape is discontinuous.
In this case a couple of alterations to the method must be made.
First, the gradients in \cref{eq:kd} are ill-defined, so $\kappa_d$ can be calculated using just the second term in the equation.
Next we must use a minimiser that does not require a differentiable optimisation function; for this we use simulated annealing \cite{Kirkpatrick1983}.
This has a chance of randomly jumping one state over the dividing barrier, but we can reduce this probability by limiting the initial temparature and maximum random displacement. We typically employ $T_0 = E_B / 10$, and $d_\text{max} = d(\bm{x}_1, \bm{x}_2) / 100$.
This has been tested using a Lennard-Jones 7-particle cluster with hard cores, resulting in a discontinuous landscape.
Using the gradient-free approach, BITSS is able to successfully find the transition states, allowing us to plot the disconnectivity graph of the system, shown in \cref{fig:flatdiscontinuous}b.
A comparison with the results for the standard Lennard-Jones cluster is given in the supplementary information \cite{Note1}.
It is worth noting, however, that a gradient-based approach is significantly more efficient, and so should be preferred if possible.

%\section{Conclusion}
In conclusion, we have demonstrated a novel method for finding transition states that tackles problems not currently addressed by existing methods.
The method is shown to be more efficient than existing double-ended methods, both in memory and computational cost, with further reductions possible by utilising adaptive remeshing techniques.
It can also be adapted to problems with complex energy landscapes, including those with flat or discontinuous regions.
This method will enable the analysis of energy landscapes for new systems in both particulate and continuum applications, and increase the feasibility of such methods for large-scale, complex systems.

\appendix
\section{}

>>>>

\documentclass[twocolumn,10pt]{revtex4-2}

@@

\section{Introduction}
Transition states are central to the description of reconfiguration mechanisms for systems in chemistry, condensed matter physics, and engineering.
Historically, many computational methods for locating transition states have grown from an atomistic or particulate perspective.
These have proven to be important tools for understanding, for example, protein folding \cite{Bryngelson1995,Onuchic1997}, biological and industrial catalysis \cite{Boehr2006,Kerns2015,Guo2018a}, quantum tunneling \cite{Richardson2016,Vaillant2019}, crystallisation \cite{Richard2018}, and cluster formation \cite{Wales1998,Wales2012}.

@@

\section{Method}
The method begins by first initialising the states, $\bm{x}_1$ and $\bm{x}_2$, in the basins of attraction of different local minima.
These can be set to the minima, but this is not a necessary requirement.
The BITSS method then leads these states to meet at the lowest point on the ridge between the basins of attraction - the transition state, as shown in \cref{fig:toy2d}a.
This is done by minimising the energy of the pair while constraining their separation and energy difference.
The separation constraint $d_i$ is incrementally reduced, such that at iteration $i$,
\begin{equation}\label{eq:diteration}
  d_i = (1 - f) d_{i-1},
\end{equation}
with $d_0$ the separation between the initial states.
A factor of $f = 0.5$ is succesful for most applications, but this can be made smaller to ensure that the states do not slide off the ridge.
The secondary constraint sets the energies of the two states equal to ensure that neither state is pulled over the ridge.

@@

\section{Results}
\subsection{Speed Comparison}
\begin{figure*}[tb]
  \includegraphics{speedtest.pdf}
  \caption{\label{fig:speedtest}
    Comparisons between BITSS, DNEB, and the string method in the convergence to the transition state as a function of the number of calculations of the energy gradient.
    Three test systems are used: (a) a Lennard-Jones seven-particle cluster, (b) cylindrical shell buckling, and (c) wetting of a chemically-striped surface.
    The configurations shown correspond to the two minimum energy states and the transition state, marked by an asterisk.
    The string method and DNEB are repeated with a differing number of images, as listed in the legend in (c).
  }
\end{figure*}
As a result of the BITSS method using only two states it has the potential to be significantly faster than chain-of-states approaches which typically use tens of states.
This is tested here by measuring the speed of convergence to the transition state for BITSS in comparison with the string method and DNEB for several representative examples.

@@

\subsection{Adaptive discretisation}
\begin{figure}[tb]
  \includegraphics{adaptivemesh.pdf}
  \caption{\label{fig:adaptivemesh}
    (a) Snapshots of the BITSS method applied to the buckling of a cylinder with a changing mesh.
        The radial displacement relative the unbuckled cylinder is shown, as well as the underlying triangular mesh.
    (b) BITSS applied to the striped wetting example with different resolutions for the two states.
        Each grid cell denotes 50x50 lattice nodes.
        The zoomed axis shows the difference in the fluid interface between the two final states, as well as the approximated transition state (solid black line).
        This is compared to the transition state found using a high resolution (dashed line).
  }
\end{figure}

@@

\subsection{Locally flat landscapes}
\begin{figure}[tb]
  \includegraphics{flatdiscontinuous.pdf}
  \caption{\label{fig:flatdiscontinuous}
    (a) Energy profile of the BITSS pathway on a 2D potential with flat regions.
        Blue and red dots denote the minima and transition state, respectively.
        Points of interest are labelled by i--iii (see text).
        Top inset: A zoomed in view around the transition state.
        Bottom inset: The pathway taken, with the edges of the flat regions marked by dashed lines.
    (b) Disconnectivity graph of the energy landscape for a seven-particle cluster with a discontinuous hard-core pair potential (orange), and a standard Lennard-Jones potnetial (gray).
        The two graphs are offset for visibility.
    (c) Plot of the discontinuous potential.
        The repulsive section of the standard Lennard-Jones potential is also shown in grey.
  }
\end{figure}

@@

\subsection{Undefined gradients}
Finally, we investigate the application of BITSS to systems with undefined gradients, such as when the landscape is discontinuous.
In this case a couple of alterations to the method must be made.
First, the gradients in \cref{eq:kd} are ill-defined, so $\kappa_d$ can be calculated using just the second term in the equation.
Next we must use a minimiser that does not require a differentiable optimisation function; for this we use simulated annealing \cite{Kirkpatrick1983}.
This has a chance of randomly jumping one state over the dividing barrier, but we can reduce this probability by limiting the initial temparature and maximum random displacement. We typically employ $T_0 = E_B / 10$, and $d_\text{max} = d(\bm{x}_1, \bm{x}_2) / 100$.
This has been tested using a Lennard-Jones 7-particle cluster with hard cores, resulting in a discontinuous landscape.
Using the gradient-free approach, BITSS is able to successfully find the transition states, allowing us to plot the disconnectivity graph of the system, shown in \cref{fig:flatdiscontinuous}b.
A comparison with the results for the standard Lennard-Jones cluster is given in the supplementary information \cite{Note1}.
It is worth noting, however, that a gradient-based approach is significantly more efficient, and so should be preferred if possible.

\section{Conclusion}
In conclusion, we have demonstrated a novel method for finding transition states that tackles problems not currently addressed by existing methods.
The method is shown to be more efficient than existing double-ended methods, both in memory and computational cost, with further reductions possible by utilising adaptive remeshing techniques.
It can also be adapted to problems with complex energy landscapes, including those with flat or discontinuous regions.
This method will enable the analysis of energy landscapes for new systems in both particulate and continuum applications, and increase the feasibility of such methods for large-scale, complex systems.

\appendix
